---
title: "WorldCLim Clipping"
author: "Dimitri Brosens"
date: "26 juni 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#R {raster}: Data acquisition – SRTM, Worldclim, Global adm. boundaries

Original Tutorial [here](https://www.gis-blog.com/r-raster-data-acquisition/)

The raster package is not only a great tool for raster processing and calculation but also very useful for data acquisition. With the function getData() you can download the following data directly into R and process it:

SRTM 90 (elevation data with 90m resolution between latitude  -60 and 60)
World Climate Data (Tmin, Tmax, Precip, BioClim)
Global adm. boundaries (different levels)

# Global adm. boundaries

## Install needed packages

```{r message=FALSE}
install.packages("raster")
library(raster)
library(sp)

```
```{r}
Belgium0 <- getData('GADM', country='BEL', level=0)
head(Belgium0)
```



* Select Dataset: The first argument specifies the dataset. ‘GADM’ returns the global administrative boundaries.
* Select country: The second argument provides the country name of the boundaries by using its ISO A3 country code [more info here](http://kirste.userpage.fu-berlin.de/diverse/doc/ISO_3166.html)
Specify level: The third argument specifies the level of of administrative subdivision (0=country, 1=first level subdivision).
The code above returns the boundaries for Belgium for the level 0. Lets compare them to the Level 1 subdivision by plotting both of them:

```{r}
#Get Data
Belgium0 <- getData('GADM' , country="BEL", level=0)
Belgium1 <- getData('GADM' , country="BEL", level=1)

#Plot
par(mfrow(2,1))
plot(Belgium0, main="Adm. Boundaries Austria Level 0")
plot(Belgium1, main="Adm. Boundaries Austria Level 1")
```


