---
title: "Clean Data for SDM"
author: "Dimitri Brosens"
date: "12 september 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#' Load libraries:

```{r}

library(tidyverse) # For data transformations
# None core tidyverse packages:
library(magrittr)  # For %<>% pipes
# Other packages
library(janitor)   # For cleaning input data
library(knitr)     # For nicer (kable) tables
library(stringr)   # For string manipulation
```

#' Set file paths (all paths should be relative to this script):
```{r}
#set RAW files path
raw_data = "../csv/papilio_machaon_SDM.csv"

#set processed files path
maxent_input_file = "../csv/processed/papilio_maxent.csv"
```

#Read raw data

```{r}
raw_data <- read.csv(file="../csv/Papilio_machaon.csv", header=TRUE, sep="\t")  #to read tab separated files
```

# Clean the data somewhat

```{r}
raw_data %<>%
  remove_empty("rows") %>%    # Remove empty rows
  clean_names()               # Have sensible (lowercase) column names
```

# Select the columns needed (decimalLatitude, decimalLongitude, scientificName)

For maxent niche modeling we need scientificName, decimalLatitude, decimalLongitude, in this order)


```{r}
raw_data %<>%
  select(scientificname, decimallatitude, decimallongitude)
  
```
#Organize column scientificName, we only need genus and speciesname

```{r}
raw_data %<>%
  mutate(scientificname = recode(scientificname,"Papilio machaon Linnaeus, 1758" = "Papilio machaon"))
```


# write data

```{r}
write.csv(raw_data, file = maxent_input_file, na = "", row.names = FALSE, fileEncoding = "UTF-8")

```

